---
title: "india_imr_1"
author: "Advait Rajagopal"
date: "June 25, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.align='center', fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```
## Data loading and munging

```{r}
#Loading packages and specs
library (dplyr)
library (rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
library(broom)
```

```{r}
#Read in data
setwd("/Users/Advait/Desktop/New_School/Fall16/BDA/Finalproject/Infant_Mortality_India")
df <- read.csv("Julian_Final_Form.csv",
               header = TRUE,
               skip = 2)
str(df)

#Drop Andaman and Nicobar & Dadra NH & Daman and Diu & Lakshadweep & Chandigarh and 2003
df <- df[df$Year != 2003,]
df <- df[df$State != "Andaman & Nicobar Islands"
         & df$State != "Dadra & Nagar Haveli"
         & df$State != "Daman & Diu"
         & df$State != "Lakshadweep"
         & df$State != "Chandigarh",]

#Fix names of Orissa, Jammu Kashmir and Chattisgarh
df$State <- as.character ( df$State )
df$State[ df$State == " Odisha "] <- " Orissa "
df$State [ df$State == " Jammu  &  Kashmir "] <- " Jammu  and  Kashmir "
df$State [ df$State == " Chhatisgarh "] <- " Chhattisgarh "
```
```{r}
#read in population data
#list.files()
pop <- read.csv("population.csv",
                header = T)

#attaching population to the data frame
ind <- merge(df,pop)
str(ind)

# Function for projecting population up or down
pop.func <- function (year, pop2011){
  pop <- NULL
  pop <- ifelse ( year > 2011 ,
                  pop2011 *(1.0184^( year - 2011)) ,
                  ifelse ( year < 2011,
                           pop2011 *((1/1.0184)^(2011 - year )),
                           pop2011 ))
  return (pop)
}
#Add column of projected population numbers
ind$population <- pop.func ( ind$Year , ind$X2011pop )
# Remove old population numbersx
ind <- ind [,-8]

##
#Put all currency in Ruppee
ind$exp.rup <- NULL
for (i in 1: length (ind [ ,1])){
  ind$exp.rup[i] <- ifelse(ind$Units [i] == "Lakh",
                           ind$Health.Exp[i]*100000 ,
                           ind$Health.Exp[i]*1000000)
}
  
#
#Get spending per capita
ind$percap <- ind$exp.rup/ ind$population

# Create ratio of female to male IMR
ind$fm.imr <- ind$Female / ind$Male

# Regress IMR ratio on health exp
summary(lm(fm.imr ~ percap ,
            data = ind ))

# Sort by year
ind <- ind[ order ( ind$Year ),]
# Create state id
ind <- within (ind , state.id <- match (State , unique ( State )))
# Create Year id
ind <- within (ind , year.id <- match (Year , unique ( Year )))
```

## Some exploratory plots
1. This plot compares IMR in 2005 and 2012 to get an idea of how IMR has moved in the specified time span. It is clear that except for Mizoram, the IMR in 2012 is less than or equal to the IMR in 2005 for all other states.

```{r, echo = FALSE}
#To view IMR across states in 2005 and 2012 ----
state <- ind$State[1:30]
c1 <- ind$Person[ind$Year == 2005]
c2 <- ind$Person[ind$Year == 2012]
plot512_df <- data.frame(state,c1,c2)
plot512_df <- plot512_df[order(plot512_df$c1, decreasing = F),]
#2005 IMR
dots_05 <- plot512_df$c1
#2012 IMR
dots_12 <- plot512_df$c2

#Plot
par(mar = c(4,6,2.5,3),
    col = "gray")
plot(dots_05,1:30,
     pch = 16,
     cex = 0.8,
     col = "white",
     yaxt = 'n',
     xaxt = 'n',
     ylab = "",
     xlab = "",
     cex.lab = 0.7,
     xlim = c(0,80),
     main = "IMR in 2005 and 2012",
     cex.main = 0.8)
abline(h = c(1:30), lty = 3)
points(dots_05,1:30,
       pch = 1,
       cex = 0.8,
       col = "black")
points(dots_12,1:30,
       pch = 16,
       cex = 0.8,
       col = "black")
axis(2, 
     at = pretty(1:30,n = 30),
     labels = paste(plot512_df$state),
     las = 2, 
     lwd.tick = 0.5,
     cex.axis = 0.7, 
     col = "gray",
     tick = T)
axis(1, 
     at = (seq(0,80,length.out = 5)),
     labels = (seq(0,80,length.out = 5)), 
     col.ticks = "gray", 
     lwd.tick = 0.5, 
     cex.axis = 0.7,
     col = "gray")

title(xlab="Infant Mortality Rate", 
      line=2, cex.lab=0.8)
legend('bottomright', 
       legend = c("2012", "2005"), 
       col = "black",
       pch = c(16,1),
       cex = 0.7,
       bty = 'n',
       text.col = "black")
#dev.off()
```

2. This plot shows the trend in IMR in the 30 Indian states under consideration for the specified time period 2005 - 2012.

```{r, echo = FALSE}
par( mfrow = c(5, 6),
     mar = c(2 , 2 , 1, 1),
     cex = 0.6 ,
     las = 1,
     col = "gray")
for (i in 1:30){
  plot ( ind$Year [ ind$state.id == i],
         ind$Person [ ind$state.id == i],
         type = "l",
         ylim = c(0, 80) ,
         yaxt = "n",
         xaxt = "n",
         xlab = "n",
         main = paste (unique(ind$State[ind$state.id == i])) ,
         cex.main = 1,
         col = "black")
  axis (1, at = c(2005 , 2012) ,
        lwd = 0.5 ,
        cex.axis = 0.8,
        col = "gray")
  axis (2, at = c(0, 80) ,
        lwd = 0.5 ,
        cex.axis = 0.6,
        col = "gray")
  title ( ylab = "IMR",
          line =0, cex.lab= .6)
  title ( xlab = "year",
          line =0, cex.lab= .6)
}
#dev.off()
```

3. This plot shows the variation in IMR for males and females by state in the specified time period. The year 2007 is where a sharp drop in female IMR is observed. In red the values of female IMR in 2006, 2007 and 2008 are shown to reflect the sharp drop and subsequent rise in female IMR.

```{r, echo = FALSE}
par( mfrow = c(5, 6),
     mar = c(2 , 2 , 1, 1),
     cex = 0.6 ,
     las = 1,
     col = "gray")
for (i in 1:30){
  plot ( ind$Year[ind$state.id == i],
         ind$Male[ind$state.id == i],
         type = "l",
         ylim = c(-5, 90) ,
         yaxt = "n",
         xaxt = "n",
         xlab = "n",
         main = paste(unique(ind$State[ind$state.id == i])) ,
         cex.main = 1,
         col = "blue",
         xlim = c(2005,2012))
  lines(ind$Year[ind$state.id == i],
        ind$Female[ind$state.id == i],
        col = "red")
  axis (1, at = c(2005 , 2012) ,
        lwd = 0.5 ,
        cex.axis = 0.8,
        col = "gray")
  axis (2, at = c(0, 80) ,
        lwd = 0.5 ,
        cex.axis = 0.6,
        col = "gray")
  title ( ylab = "IMR",
          line =0, cex.lab= .6)
  title ( xlab = "year",
          line =0, cex.lab= .6)
  text(2007, 
       ind$Female[ind$state.id == i & ind$year.id == 3],
       labels = ind$Female[ind$state.id == i & ind$year.id == 3],
       pos = 1,
       cex = 0.7,
       col = "red")
  text(2006, 
       ind$Female[ind$state.id == i & ind$year.id == 2],
       labels = ind$Female[ind$state.id == i & ind$year.id == 2],
       pos = 3,
       cex = 0.7,
       col = "red")
  text(2008, 
       ind$Female[ind$state.id == i & ind$year.id == 4],
       labels = ind$Female[ind$state.id == i & ind$year.id == 4],
       pos = 3,
       cex = 0.7,
       col = "red")
}
```

4. This plot is the ratio of female to male IMR. The dotted line is at one. If the line is above the dotted line, more female infants die and if the line is below the dotted line more males die with respect to females.

```{r, echo = FALSE}
par( mfrow = c(5, 6),
     mar = c(2 , 2 , 1, 1),
     cex = 0.6 ,
     las = 1,
     col = "gray")
for (i in 1:30){
  plot ( ind$Year [ ind$state.id == i],
         ind$fm.imr[ ind$state.id == i],
         type = "l",
         ylim = c(0.4, 2.5) ,
         yaxt = "n",
         xaxt = "n",
         xlab = "n",
         main = paste ( unique ( ind$State [ ind$state.id == i])) ,
         cex.main = 1,
         col = "black")
  axis (1, at = c(2005 , 2012) ,
        lwd = 0.5 ,
        cex.axis = 0.8,
        col = "gray")
  axis (2, at = c(0.4,1,2.5) ,
        lwd = 0.5 ,
        cex.axis = 0.6,
        col = "gray")
  title ( ylab = "IMR Ratio",
          line =0, cex.lab= .6)
  title ( xlab = "year",
          line =0, cex.lab= .6)
  abline(h = 1, lty = 2)
}
```

## Modeling 
I model IMR as a function of percapita health expenditure using hierarchical Bayesian regression. I use health expenditure lagged by one period. I have a state level intercept, a time intercept and state level coefficients for the impact of health expenditure on IMR. 

```{r}
#Some data munging to get the data ready for Stan. Also creating lags here.

###Re order by state id for lags to work
#creating "ind_temp" so as to not mess with the original data frame "ind"
ind_temp <- ind[order(ind$state.id),]
ind_temp <- ind_temp %>%
  group_by(State) %>%
  mutate(lag.exp = lag(percap, order_by = State))
#check if it works
# cbind(ind_temp$State, 
#       ind_temp$Year, 
#       ind_temp$year.id, 
#       ind_temp$percap, 
#       ind_temp$lag.exp)
#turns out it does, now add ind_temp$lag.exp to ind
ind <- ind[order(ind$state.id),]
ind$lag.exp <- ind_temp$lag.exp
# ind
#Drop NA in Year 2005
ind <- ind[complete.cases(ind),]
str(ind)

###Preparing variables for Stan Model
#years start at 2006 so re-code year.id
year_id <- ind$year.id - 1
state_id <- ind$state.id
# imr <- ind$Person
imr_ratio <- ind$fm.imr
percap_exp <- ind$lag.exp
N <- length(ind$Person)
```

# Model 1 - The Normal Model

log(IMR(it)) ~ N(alpha_s(i) + alpha_t(t) + beta_exp(i) * log(percapita_exp(it)), sigma^2)

```{r, results="hide"}
#Model 1
#running Model 1 ----
#first run model_1 for females and save results
imr <- ind$Female
stanc("model_1.stan")
fit_1_f <- stan("model_1.stan",
              data = list("N",
                          "year_id",
                          "state_id",
                          "imr",
                          "percap_exp"),
              iter = 2000,
              chains = 4)
#second run model_1 for males and save results
imr <- ind$Male
stanc("model_1.stan")
fit_1_m <- stan("model_1.stan",
                data = list("N",
                            "year_id",
                            "state_id",
                            "imr",
                            "percap_exp"),
                iter = 2000,
                chains = 4)

#Preparing plots for Model 1
#extracting the results into dataframes
tidy_1_f <- tidy(fit_1_f, conf.int = T, conf.level = 0.50)
tidy_1_m <- tidy(fit_1_m, conf.int = T, conf.level = 0.50)
#
female_pred_1 <- tidy_1_f[75:284,2]
female_low_1 <- tidy_1_f[75:284,4]
female_hi_1 <- tidy_1_f[75:284,5]
#
male_pred_1 <- tidy_1_m[75:284,2]
male_low_1 <- tidy_1_m[75:284,4]
male_hi_1 <- tidy_1_m[75:284,5]
```

Viewing the results of the modeling.

1. The state intercepts

```{r, echo = FALSE}
#State intercepts for model 1
par(mar = c(2,6,2,1),
    mfcol = c(2,1),
    col = "gray")
plot(exp(tidy_1_f[1:30,2]),
     seq(1,30),
     pch = 16,
     col = "black",
     main = "State intercepts for Females - Model 1",
     cex.main = 0.6,
     xlab = NA,
     xaxt = 'n',
     ylab = NA,
     yaxt = 'n',
     xlim = c(0,40),
     cex = 0.7)
abline (h = c(1:30), lty = 3)
arrows(exp(tidy_1_f[1:30,2]),
       seq(1,30),
       exp(tidy_1_f[1:30,4]),
       seq(1,30),
       col = "gray",
       length = 0,
       lwd = 2)
arrows(exp(tidy_1_f[1:30,2]),
       seq(1,30),
       exp(tidy_1_f[1:30,5]),
       seq(1,30),
       col = "gray",
       length = 0,
       lwd = 2)
axis(2, 
     at = pretty(1:30,n = 30),
     labels = paste(unique(ind$State)),
     las = 2, 
     lwd.tick = 0.5,
     cex.axis = 0.3, 
     col = "gray",
     tick = T)
axis (1, seq(0,40, length.out = 5) ,
      lwd = 0.5 ,
      cex.axis = 0.6,
      col = "gray")
points(exp(tidy_1_f[1:30,2]),
       seq(1,30),
       pch = 16,
       col = "black",
       cex = 0.7)
legend ('bottomright',
        legend = c("Data", "50% Interval"),
        col = c("black  " , "gray") ,
        pch=c(16,3),
        lty = c(0, 1),
        # bty = 'n',
        cex = 0.5,
        text.col = "black")
plot(exp(tidy_1_m[1:30,2]),
     seq(1,30),
     pch = 16,
     col = "black",
     main = "State intercepts for Males - Model 1",
     cex.main = 0.6,
     xlab = NA,
     xaxt = 'n',
     ylab = NA,
     yaxt = 'n',
     xlim = c(0,40),
     cex = 0.7)
abline (h = c(1:30), lty = 3)
arrows(exp(tidy_1_m[1:30,2]),
       seq(1,30),
       exp(tidy_1_m[1:30,4]),
       seq(1,30),
       col = "gray",
       length = 0,
       lwd = 2)
arrows(exp(tidy_1_m[1:30,2]),
       seq(1,30),
       exp(tidy_1_m[1:30,5]),
       seq(1,30),
       col = "gray",
       length = 0,
       lwd = 2)
axis(2, 
     at = pretty(1:30,n = 30),
     labels = paste(unique(ind$State)),
     las = 2, 
     lwd.tick = 0.5,
     cex.axis = 0.3, 
     col = "gray",
     tick = T)
axis (1, seq(0,40, length.out = 5) ,
      lwd = 0.5 ,
      cex.axis = 0.6,
      col = "gray")
points(exp(tidy_1_m[1:30,2]),
       seq(1,30),
       pch = 16,
       col = "black",
       cex = 0.7)
legend ('bottomright',
        legend = c("Data", "50% Interval"),
        col = c("black  " , "gray") ,
        pch=c(16,3),
        lty = c(0, 1),
        # bty = 'n',
        cex = 0.5,
        text.col = "black")

```

2. The time intercepts
```{r, echo  = FALSE}
#time intercept for model 1 ----
par(mar = c(2.5,2.4,2,2),
    mfcol = c(1,2),
    col = "gray")
plot(exp(tidy_1_f[33:39,2]),
     seq(2006,2012),
     pch = 16,
     col = "black",
     main = "Time intercepts for Females - Model 1",
     cex.main = 0.8,
     xlab = NA,
     xaxt = 'n',
     ylab = NA,
     yaxt = 'n',
     xlim = c(0,20))
abline (h = c(2006:2012), lty = 3)
arrows(exp(tidy_1_f[33:39,2]),
       seq(2006,2012),
       exp(tidy_1_f[33:39,4]),
       seq(2006,2012),
       col = "gray",
       length = 0,
       lwd = 2)
arrows(exp(tidy_1_f[33:39,2]),
       seq(2006,2012),
       exp(tidy_1_f[33:39,5]),
       seq(2006,2012),
       col = "gray",
       length = 0,
       lwd = 2)
axis (2, at = seq(2006 , 2012) ,
      lwd = 0.5 ,
      cex.axis = 0.6,
      col = "gray",
      las = 2)
axis (1, seq(0,20, length.out = 5) ,
      lwd = 0.5 ,
      cex.axis = 0.6,
      col = "gray")
points(exp(tidy_1_f[33:39,2]),
       seq(2006,2012),
       pch = 16,
       col = "black")
legend (14,2011.8,
        legend = c("Data", "50% Interval"),
        col = c("black  " , "gray") ,
        pch=c(16,3),
        lty = c(0, 1),
        bty = 'n',
        cex = 0.5,
        text.col = "black")
plot(exp(tidy_1_m[33:39,2]),
     seq(2006,2012),
     pch = 16,
     col = "black",
     main = "Time intercepts for Males - Model 1",
     cex.main = 0.8,
     xlab = NA,
     xaxt = 'n',
     ylab = NA,
     yaxt = 'n',
     xlim = c(0,20))
abline (h = c(2006:2012), lty = 3)
arrows(exp(tidy_1_m[33:39,2]),
       seq(2006,2012),
       exp(tidy_1_m[33:39,4]),
       seq(2006,2012),
       col = "gray",
       length = 0,
       lwd = 2)
arrows(exp(tidy_1_m[33:39,2]),
       seq(2006,2012),
       exp(tidy_1_m[33:39,5]),
       seq(2006,2012),
       col = "gray",
       length = 0,
       lwd = 2)
axis (2, at = seq(2006, 2012) ,
      lwd = 0.5 ,
      cex.axis = 0.6,
      col = "gray",
      las = 2)
axis (1, seq(0,20, length.out = 5) ,
      lwd = 0.5 ,
      cex.axis = 0.6,
      col = "gray")
points(exp(tidy_1_m[33:39,2]),
       seq(2006,2012),
       pch = 16,
       col = "black")
legend (14,2011.8,
        legend = c("Data", "50% Interval"),
        col = c("black  " , "gray") ,
        pch=c(16,3),
        lty = c(0, 1),
        bty = 'n',
        cex = 0.5,
        text.col = "black")
```

3. The marginal posterior of the beta or slope coefficient.
```{r, echo = FALSE}
ext_1_f <- extract(fit_1_f)
ext_1_m <- extract(fit_1_m)
par( mfrow = c(5, 6),
     mar = c(2 , 1 , 1, 1),
     cex = 0.6 ,
     las = 1,
     col = "gray")
for(i in 1:30){
  plot(density(ext_1_f$beta_exp[,i]),
       main = paste ( unique ( ind$State [ ind$state.id == i])),
       cex.main = 0.8,
       xlab = NA,
       ylab = NA,
       xaxt = 'n',
       yaxt = 'n',
       col = "red",
       xlim = c(-0.35,0.20),
       bty = 'n')
  lines(density(ext_1_m$beta_exp[,i]),
        col = "blue")
  abline(v = 0, lty = 2)
  axis (1, at = c(-0.35, 0, 0.20) ,
        lwd = 0.5 ,
        cex.axis = 0.6,
        col = "gray")
  legend ('topright', legend = c("F", "M"),
          col = c("red" , "blue"),
          lty = c(1, 1),
          bty = 'n',
          cex = 0.5,
          text.col = "black")
}
```

Model 1 - Posterior Predictive Checking for Females and Males. 

The horizontal axes show the predicted values and the vertical axes show the original or true vales. The values cluster around the 45 degree line showing that the predictions are pretty good. The gray lines around the points show the 50% interval.
```{r, echo = FALSE}
#female posterior predictive check ----
par(mfcol = c(1,2),
    col = "gray")
plot(exp(female_pred_1), 
     ind$Female,
     pch = 16,
     cex = 0.7,
     col = "black",
     xaxt = 'n',
     xlab = "predicted female IMR",
     yaxt = 'n',
     ylab = "actual female IMR",
     cex.lab = 0.8,
     main = "Model 1 - Posterior Predictive Checks - Female IMR",
     cex.main = 0.6,
     xlim = c(0,80),
     ylim = c(0,80))
axis (1, at = seq(0,80, length.out = 9),
      lwd = 0.5 ,
      cex.axis = 0.8,
      col = "gray")
axis (2, at = seq(0,80, length.out = 9) ,
      lwd = 0.5 ,
      cex.axis = 0.8,
      col = "gray")
abline(0,1,
       col = "black",
       lty = 1)
arrows(exp(female_pred_1),
       ind$Female,
       exp(female_low_1),
       ind$Female,
       col = "gray",
       length = 0)
arrows(exp(female_pred_1),
       ind$Female,
       exp(female_hi_1),
       ind$Female,
       col = "gray",
       length = 0)
points(exp(female_pred_1), 
       ind$Female,
       pch = 16,
       cex = 0.7,
       col = "black")
legend ('topleft', legend = c("Data", "50% Interval"),
        col = c("black  " , "darkgrey") , pch=c(16,3),
        lty = c(0, 1),
        bty = 'n',
        cex = 0.7,
        text.col = "black")
#Males
plot(exp(male_pred_1), 
     ind$Male,
     pch = 16,
     cex = 0.7,
     col = "black",
     xaxt = 'n',
     xlab = "predicted male IMR",
     yaxt = 'n',
     ylab = "actual male IMR",
     cex.lab = 0.8,
     main = "Model 1 - Posterior Predictive Checks - Male IMR",
     cex.main = 0.6,
     xlim = c(0,80),
     ylim = c(0,80))
axis (1, at = seq(0,80, length.out = 9),
      lwd = 0.5 ,
      cex.axis = 0.8,
      col = "gray")
axis (2, at = seq(0,80, length.out = 9) ,
      lwd = 0.5 ,
      cex.axis = 0.8,
      col = "gray")
abline(0,1,
       col = "black",
       lty = 1)
arrows(exp(male_pred_1),
       ind$Male,
       exp(male_low_1),
       ind$Male,
       col = "gray",
       length = 0)
arrows(exp(male_pred_1),
       ind$Male,
       exp(male_hi_1),
       ind$Male,
       col = "gray",
       length = 0)
points(exp(male_pred_1), 
       ind$Male,
       pch = 16,
       cex = 0.7,
       col = "black")
legend ('topleft', legend = c("Data", "50% Interval"),
        col = c("black  " , "darkgrey") , pch=c(16,3),
        lty = c(0, 1),
        bty = 'n',
        cex = 0.7,
        text.col = "black")
```


Posterior predictive checks in a grid for model 1. The following two plots show the performance of the model alongside the true data. The data is in black. The gray solid lines represent the prediction and the gray dashed lines represent the 50% posterior intervals.

1. Females

```{r, echo = FALSE}
ppc <- data.frame(ind$State,
                  ind$state.id,
                  ind$Year,
                  female_pred_1,
                  female_low_1,
                  female_hi_1,
                  male_pred_1,
                  male_low_1,
                  male_hi_1)
ppc$female <- ind$Female
ppc$male <- ind$Male

#Female grid PPC - model 1
par( mfrow = c(5, 6),
     mar = c(2 , 2 , 1, 1),
     cex = 0.6 ,
     las = 1,
     col = "gray")
for (i in 1:30){
  plot ( ppc$ind.Year[ppc$ind.state.id == i],
         ppc$female[ppc$ind.state.id == i],
         type = "l",
         ylim = c(0, 80) ,
         yaxt = "n",
         xaxt = "n",
         xlab = "n",
         main = paste ( unique ( ppc$ind.State[ ppc$ind.state.id == i])) ,
         cex.main = 1,
         col = "black")
  lines(ppc$ind.Year[ppc$ind.state.id == i],
        exp(ppc$female_pred_1[ppc$ind.state.id == i]),
        col = "gray")
  lines(ppc$ind.Year[ppc$ind.state.id == i],
        exp(ppc$female_low_1[ppc$ind.state.id == i]),
        col = "gray",
        lty = 2)
  lines(ppc$ind.Year[ppc$ind.state.id == i],
        exp(ppc$female_hi_1[ppc$ind.state.id == i]),
        col = "gray",
        lty = 2)
  axis (1, at = c(2006 , 2012) ,
        lwd = 0.5 ,
        cex.axis = 0.8,
        col = "gray")
  axis (2, at = c(0, 80) ,
        lwd = 0.5 ,
        cex.axis = 0.6,
        col = "gray")
  title ( ylab = "IMR",
          line =0, cex.lab= .6)
  title ( xlab = "year",
          line =0, cex.lab= .6)
}
```

2. Males

```{r, echo = FALSE}
#Male grid PPC - model 1 
par( mfrow = c(5, 6),
     mar = c(2 , 2 , 1, 1),
     cex = 0.6 ,
     las = 1,
     col = "gray")
for (i in 1:30){
  plot ( ppc$ind.Year[ppc$ind.state.id == i],
         ppc$male[ppc$ind.state.id == i],
         type = "l",
         ylim = c(0, 80) ,
         yaxt = "n",
         xaxt = "n",
         xlab = "n",
         main = paste ( unique ( ppc$ind.State[ ppc$ind.state.id == i])) ,
         cex.main = 1,
         col = "black")
  lines(ppc$ind.Year[ppc$ind.state.id == i],
        exp(ppc$male_pred_1[ppc$ind.state.id == i]),
        col = "gray")
  lines(ppc$ind.Year[ppc$ind.state.id == i],
        exp(ppc$male_low_1[ppc$ind.state.id == i]),
        col = "gray",
        lty = 2)
  lines(ppc$ind.Year[ppc$ind.state.id == i],
        exp(ppc$male_hi_1[ppc$ind.state.id == i]),
        col = "gray",
        lty = 2)
  axis (1, at = c(2006 , 2012) ,
        lwd = 0.5 ,
        cex.axis = 0.8,
        col = "gray")
  axis (2, at = c(0, 80) ,
        lwd = 0.5 ,
        cex.axis = 0.6,
        col = "gray")
  title ( ylab = "IMR",
          line =0, cex.lab= .6)
  title ( xlab = "year",
          line =0, cex.lab= .6)
}
```


<!-- # Model 2 - The Gamma Model -->
<!-- I model the ratio of female to male IMR with a Gamma distribution, whose shape parameters are functions of the per capita health expenditure. -->

<!-- ```{r, results = "hide"} -->
<!-- #Model 2 ---- -->
<!-- stanc("model_2.stan") -->
<!-- imr <- ind$fm.imr -->
<!-- fit_2 <- stan("model_2.stan", -->
<!--                   data = list("N", -->
<!--                               "imr", -->
<!--                               "percap_exp", -->
<!--                               "state_id", -->
<!--                               "year_id"), -->
<!--                   iter = 2000, -->
<!--                   chains = 4) -->
<!-- print(fit_2) -->
<!-- gamma_fit <- tidy(fit_2,conf.int = T, conf.level = 0.50 )  -->
<!-- str(gamma_fit) -->
<!-- imr_ratio_pred <- gamma_fit[425:634,2] -->
<!-- imr_ratio_low <- gamma_fit[425:634,4] -->
<!-- imr_ratio_hi <- gamma_fit[425:634,5] -->
<!-- ``` -->

<!-- Model 2 - Posterior predictive checks -->
<!-- ```{r, echo = FALSE} -->
<!-- x <- c(1:210) -->
<!-- par(col = "gray") -->
<!-- plot(imr, -->
<!--      pch = 16, -->
<!--      cex = 0.7, -->
<!--      ylim = c(0,5), -->
<!--      yaxt = 'n', -->
<!--      xaxt = 'n', -->
<!--      ylab = "Ratio of Female to Male IMR", -->
<!--      cex.lab = 0.7, -->
<!--      main = "Model 2 - Posterior Predictive Checks", -->
<!--      cex.main = 0.8) -->
<!-- arrows(x, -->
<!--        imr_ratio_pred, -->
<!--        x, -->
<!--        imr_ratio_low, -->
<!--        length = 0, -->
<!--        col = "gray") -->
<!-- arrows(x, -->
<!--        imr_ratio_pred, -->
<!--        x, -->
<!--        imr_ratio_low, -->
<!--        length = 0, -->
<!--        col = "gray") -->
<!-- arrows(x, -->
<!--        imr_ratio_pred, -->
<!--        x, -->
<!--        imr_ratio_hi, -->
<!--        length = 0, -->
<!--        col = "gray") -->
<!-- points(imr, -->
<!--        pch = 16, -->
<!--        cex = 0.7, -->
<!--        col = "black") -->
<!-- points(imr_ratio_pred, -->
<!--        pch = 1, -->
<!--        cex = 0.7, -->
<!--        col = "black") -->
<!-- axis (1, at = seq(0,200, length.out = 5), -->
<!--       lwd = 0.5 , -->
<!--       cex.axis = 0.8, -->
<!--       col = "gray") -->
<!-- axis (2, at = seq(0,5, length.out = 6) , -->
<!--       lwd = 0.5 , -->
<!--       cex.axis = 0.8, -->
<!--       col = "gray") -->
<!-- legend ('topleft',  -->
<!--         legend = c("Data","Prediction", "50% Interval"), -->
<!--         col = c("black","black", "darkgrey") ,  -->
<!--         pch=c(16,1,3), -->
<!--         lty = c(0,0, 1), -->
<!--         bty = 'n', -->
<!--         cex = 0.7, -->
<!--         text.col = "black") -->
<!-- ``` -->














Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
